---
title: "Machine Learnning Project"
author: "Cheng Zhiyuan"
date: "January 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

This paper is to try the different machine learning method and find a best to predict classification. The first part is use rpart method. And second part is random forest method.


```{r}
# setup the work directory and load necessary library
setwd("c:/MyWorkspace/MyLearning/Data Science/Data Science Coursera/machine learning/Project")
library(caret)
library(rattle)
library(rpart)
library(rpart.plot)
library(randomForest)
library(repmis)
```

## Process the data

```{r}
# loading the training and testing data, and remove NA data in the training set
training <- read.csv("pml-training.csv", na.strings = c("NA", ""))
training <- training[, colSums(is.na(training)) == 0]
testing <- read.csv("pml-testing.csv", na.strings = c("NA", ""))
```

```{r}
# remove the unneccesary data
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
# remove the NA column
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
```

```{r}
# split the training dataset into training and validate dataset
set.seed(198156)
trainingsetting <- createDataPartition(training$classe, p = 0.6, list = FALSE)
trainingdata <- training[trainingsetting,]
validatedate <- training[-trainingsetting,] 
```

## rpart Prediction

```{r}
# use rpart to training the model
control <- trainControl(method = "cv", number = 5)
fit_rpart = rpart(classe ~ ., data = trainingdata, method = "class")
fancyRpartPlot(fit_rpart)
```

```{r}
# use the validate dataset to verify the fit model
predict_rpart <- predict(fit_rpart,validatedate,type ="class")
confusionMatrix(validatedate$classe, predict_rpart)
```

The overall accurary of rpart method is 0.7539. 

## Ramdom Forest Prediction

```{r}
fit_rf <- train(classe ~ ., data = trainingdata, method = "rf", trControl = control)
print(fit_rf, digits = 4)
```

```{r}
# predict outcomes using validation set
predict_rf <- predict(fit_rf, validatedate)
confusionMatrix(validatedate$classe, predict_rf)
```

The overall accuracy of ramdom forest method is 0.9925

## Prediction on testing set

Now we will use ramdom forest model to predict the testing data set
```{r}
predict(fit_rf, testing)
```